Automated iterative forecasting for the Portal Project

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Cleanup edits in Discussion

We designed this forecasting system with the goal of making it relatively easy to build, maintain, and extend. We used existing technology for both running the pipeline and building individual components, which allowed is to build the system relatively cheaply in terms of both time and money. This included the use of tools like Docker for reproducibility, the Travis CI continuous integration system for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts. By using this "continuous analysis" approach [@beaulieu2017], where analyses are automatically rerun when changes are made to data, models, or associated code, we have reduced the time required by scientists to run and maintain the forecasting pipeline. To make the system extensible so that new models could be easily incorporated, we used a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 3). This should substantially lower the barriers to other scientists contributing models to this forecasting effort. We also automatically archive the resulting forecasts publicly so thatthe performance of these forecasts can be assessed by both us and other researchers as new data is collected. This serves as a form of pre-registration of expected results () by providing a quantitative record of the forecast before the data being predicted were collected.

While building this system was facilitated by the use of existing technological solutions, there were still a number of challenges in making existing tools work for automated iterative forecasting. Continuous integration is designed primarily for running automated tests on software, not for running a coordinated forecasting pipeline. As a result, extra effort was sometimes necessary to figure out how to get these systems to work properly in non-standard situations, like running code that was not part of a software package. In addition, hosted continuous integration solutions, like Travis, provide only limited computational resources. As the number and complexity of the models we fit has grown, we have had to continually invest effort in reducing our total compute time so we can stay within these limits. Finally, we found no satisfactory existing solution for archiving our results. All approaches we tried had limitations when it came to automatically generating publicly versioned archives of forecasts on a repeated basis. Overall, we found existing technology to be sufficient to the task, but it required greater expertise and a greater investment of time than is ideal. Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is possible for scientists to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems. Forecasting systems such as ours can begin the process of making iterative forecasts while we wait for more integrated and easier to deploy platforms for iterative forecasting to be developed. Once more integrated platforms exist, it will be easier to convert current systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].

@@

We developed this infrastructure for automatically making iterative forecasts with the goals of making accurate forecasts for this well-studied system, learning what methods work well for ecological forecasting more generally, and improving our understanding of the processes driving ecological dynamics. The most obvious application of automated iterative ecological forecasting is for speeding up development of forecasting models by providing the most recent data available to models and by quickly iterating to improve the models used for forecasting. By learning what works best for forecasting in this and other ecological systems, we will better understand what the best approaches are for ecological forecasting more generally. By designing the pipeline so that it can forecast many different aspects of the ecological community, we also hope to learn about what aspects of ecology are more forecastable. Finally, automated forecasting infrastructures like this one also provide a core foundation for faster scientific inquiry more broadly because new models can quickly be applied to data and compared to existing models. The forecasting infrastructure does the time-consuming work of data processing, data integration, and model assessment, allowing new research to focus on the models being developed and the inferences about the system that can be drawn from them [@dietze2018]. We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole at this long-term study site. By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@carpenter example]. By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, our understanding of the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.

>>>>

We designed this forecasting system with the goal of making it relatively easy to build, maintain, and extend. We used existing technology for both running the pipeline and building individual components, which allowed us to build the system relatively cheaply in terms of both time and money. This included the use of tools like Docker for reproducibility, the Travis CI continuous integration system for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts. By using this "continuous analysis" approach [@beaulieu2017], where analyses are automatically rerun when changes are made to data, models, or associated code, we have reduced the time required by scientists to run and maintain the forecasting pipeline. To make the system extensible so that new models could be easily incorporated, we used a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 3). This should substantially lower the barriers to other scientists contributing models to this forecasting effort. We also automatically archive the resulting forecasts publicly so that the performance of these forecasts can be assessed by both us and other researchers as new data is collected. This serves as a form of pre-registration by providing a quantitative record of the forecast before the data being predicted were collected.

While building this system was facilitated by the use of existing technological solutions, there were still a number of challenges in making existing tools work for automated iterative forecasting. Continuous integration is designed primarily for running automated tests on software, not for running a coordinated forecasting pipeline. As a result, extra effort was sometimes necessary to figure out how to get these systems to work properly in non-standard situations, like running code that was not part of a software package. In addition, hosted continuous integration solutions, like Travis, provide only limited computational resources. As the number and complexity of the models we fit has grown, we have had to continually invest effort in reducing our total compute time so we can stay within these limits. Finally, we found no satisfactory existing solution for archiving our results. All approaches we tried had limitations when it came to automatically generating publicly versioned archives of forecasts on a repeated basis. Overall, we found existing technology to be sufficient to the task, but it required greater expertise and a greater investment of time than is ideal. Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is possible for scientists to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems. Developing systems that  produce automated iterative forecasts will make it easier to convert research focused forecasts into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].

@@

We developed this infrastructure for automatically making iterative forecasts with the goals of making accurate forecasts for this well-studied system, learning what methods work well for ecological forecasting more generally, and improving our understanding of the processes driving ecological dynamics. The most obvious application of automated iterative ecological forecasting is for speeding up development of forecasting models by providing the most recent data available to models and by quickly iterating to improve the models used for forecasting. By learning what works best for forecasting in this and other ecological systems, we will better understand what the best approaches are for ecological forecasting more generally. By designing the pipeline so that it can forecast many different aspects of the ecological community, we also hope to learn about what aspects of ecology are more forecastable. Finally, automated forecasting infrastructures like this one also provide a core foundation for faster scientific inquiry more broadly because new models can quickly be applied to data and compared to existing models. The forecasting infrastructure does the time-consuming work of data processing, data integration, and model assessment, allowing new research to focus on the models being developed and the inferences about the system that can be drawn from them [@dietze2018]. We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole. By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@carpenter example]. By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, our understanding of the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.
